---
title: "Leprosy Cases"
subtitle: "Homework 2 SDS"
output: 
  html_document
date: "2024-09-19"
author: Mika Ohayon- SDS 313 UT Austin
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


# *Dataset information*
This is me creating an edit in this rmd. 
 *The data set contains values pertaining to the number of new cases of leprosy in 2022, population, GDP (billions of US \$), land area and the region.*

```{r, echo=FALSE}

lep <- read.csv("Homework2_leprosy.csv")
```

# **1: Country with the highest # of leprosy cases in 2022**

India had the largest amount of leprosy cases in 2022 with 103819 cases. However, it is not a fair comparison to look purely at the number of cases when comparing prevalence of the disease across different countries as India has the largest population overall therefore it will have a greater number amount of cases rather than the greatest proportion of cases.

```{r, echo=FALSE, results='hide'}
max(lep$New_Cases, na.rm=TRUE)
lep[which(lep$New_Cases==max(lep$New_Cases, na.rm=TRUE)),]
```

# **2: Leprosy cases per 100K people in each country**

The graph below shows the distribution of leprosy cases per 100k people in each country.
```{r, echo=FALSE, results='hide'}
casesPer100 <- (lep$New_Cases / (lep$Population/100000))
lep$casesPer100<- casesPer100

library(ggplot2)

lep_ggplot <- ggplot(lep)

lep_ggplot + geom_histogram(aes(x=casesPer100), binswidth=30)+labs(title='Frequency of Leprosy Cases Per 100K people', x="Leprosy Cases Per 100K", y= 'Frequency')

median100<- round(median(casesPer100, na.rm=TRUE), 2)
IQR<-IQR(casesPer100, na.rm=TRUE)
```

The graph portraying the amount of cases per 100K people shows a strong skew to the right, implying that there is a higher frequenycy of low numbers of cases per 100k people with median rounding to rounding to  `r median100` and an Interquartile range of `r IQR`

# **3 Cases per 100K across the different world regions**

```{r, echo=FALSE, results='hide'}
library(dplyr)
library(readr)
lep_ggplot+ geom_histogram (aes(x=casesPer100))+ facet_wrap(~Region)

```

The following table displays a summary of leprosy cases in prospective of the different world regions. The table shows ,median number of cases as well as the number of countries within each region.

The graphs all show a right skew with a higher frequency in the low number of cases. According to the graph and taable, Africa has the highest median of leprosy cases, followed by Asia/pacific , then the Americas, then the middle east, with europe having the lowest median of leprosy cases.

```{r, echo=FALSE}

new_lep<-lep%>%
  group_by(Region)%>%
  summarise(
    num_countries=n(),
    median_cases=median(casesPer100, na.rm=TRUE)
  )
new_lep

```

# **4: Is the amount of Leprosy cases/ 100K people related to a countries GDP(US Dollars)?**

The relationship between leprosy cases leprosy cases per 100k people and the GDP (US Dollars) of a country is very low as the correlation coefficient is `r round(cor(lep$casesPer100, lep$GDP_billions, use="complete.obs"), digits=2)` . The correlation coefficient is very small and rounds to 0, therefore indicating that there is close to no relationship between GDP and cases per 100K people.

```{r, echo=FALSE}
cc<-round(cor(lep$casesPer100, lep$GDP_billions, use="complete.obs"), digits=2)

lep_ggplot + geom_point(aes(x=casesPer100, y=GDP_billions)) + labs(x="Cases per 100K", y = "GDP (Billions)", title = "Relationship Between Cases and GDP")


```

# **5: Conclusion**

Through my analysis I found that countries with a higher population are likely to have more leprosy cases than those with lower populations. Specifically countries within Asia/pacific and Africa have the highest frequency of cases than any other place in the world. I also discovered that GDP (US Dollars) does not correlate to the number of cases of leprosy per country which was interesting. For more information regarding leprosy, you may visit [International Leprosy Association](https://www.bing.com/ck/a?!&&p=ab0a608bc1406b77JmltdHM9MTcyNjc5MDQwMCZpZ3VpZD0wODEwMTkyYi0yZDcxLTZhMmItMTUwZS0wYmEwMmNmZjZiZGMmaW5zaWQ9NTE5NA&ptn=3&ver=2&hsh=3&fclid=0810192b-2d71-6a2b-150e-0ba02cff6bdc&psq=International+Leprosy+Association&u=a1aHR0cHM6Ly93d3cuaW50ZXJuYXRpb25hbGxlcHJvc3lhc3NvY2lhdGlvbi5vcmcv&ntb=1).
